Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2018-03-29 revised_prop_letter_ajhs.tex
revised_prop_letter_ajhs.tex Eliminating pdfLaTeX warning...
revised_prop_letter_ajhs.tex Updated what Andy, Amr, and I discussed...

We now give the necessary technical definitions to prove this
non-empty core property.
\begin{definition}[Consistency]\label{def:Consistency} We say a QIVPM
$\bar{\mu}$ is
\emph{consistent} with a state $\rho$ and a projector~$P$ if the
interval $\bar{\mu}(P)$ contains the exact probability calculated by
the Born rule~\cite{Born1983bibTeX,Mermin2007,Jaeger2007}, i.e.,
  \begin{equation}
  \muB_{\rho}\left(P\right) = \Tr\left(\rho
  P\right)\in\bar{\mu}\left(P\right)\ .\label{eq:consistent}
  \end{equation}
In contrast with classical probability spaces, there is \emph{no
guarantee} that there exists a unique state $\rho$ that satisfies
Eq.~(\ref{eq:consistent}) and therefor{\color{OliveGreen}e} is consistent with a QIVPM.
\end{definition}

@@

{\color{red}An outline of}The proof, detailed in the forthcoming
thesis~\cite{TaiThesis2018}, proceeds as
follows. From a subspace $\eventsC$ of mutually commuting events, one
can construct a partial orthonormal basis by diagonalization, and
complete this to a full orthonormal basis $\overline{\eventsC}$.  We
can then build a bijection between the QIVPM on the set of projectors
associated with this basis and the set of classical events
corresponding to this basis.  Using this correspondence together with
the classical result by
Shapley~\cite{Shapley1971,GilboaSchmeidler1994,NgMoYeh1997,Grabisch2016},
we can establish that for the special case of commuting events, a
QIVPM will always have a non-empty core.

@@

\subsection{\color{red}Finite-Precision Extension of Gleason's Theorem}

@@

\subsection{\color{red}Ultramodular Functions}

Theorem~\ref{thm:Finite-precision-Gleason} generalizes Gleason's
theorem in the sense that it accounts for a larger class of probability measures that includes
the conventional one as a limit. The theorem is however ``special'' in
the sense that it only applies to the particular class of QIVPMs
constructed by composing a{\color{blue}n} interval map with a conventional
quantum probability measure. QIVPMs constructed in this manner have
some peculiar properties that we examine next.

A{\color{blue}n} interval map is called \emph{ultramodular} if it satisfies
the following properties.

\begin{definition}[Ultramodular Functions]\label{def:THOS}Given
  a collection of intervals $\mathscr{I}$ including $\imposs$ and
  $\necess$, a{\color{blue}n} interval map
  $\ultramodular:\left[0,1\right]\rightarrow\mathscr{I}$ is called
  ultramodular if:
\begin{subequations}\label{eq:iota-constraints}
\begin{eqnarray}
 &  & \ultramodular(0)=\imposs\,,\\
 &  & \ultramodular(1)=\necess\,,\\
 &  & \ultramodular\left(1-x\right)=\necess-\ultramodular\left(x\right)\,,
\end{eqnarray}
\end{subequations}and for any three numbers~$x_{0}$, $x_{1}$, and
$x_{2}\in\left[0,1\right]$ such that
$y=x_{0}+x_{1}+x_{2}\in\left[0,1\right]$, we have:
\begin{equation}
\ultramodular\left(y\right)+\ultramodular\left(x_{2}\right)\subseteq\ultramodular\left(x_{0}+x_{2}\right)+\ultramodular\left(x_{1}+x_{2}\right)\,.\label{eq:iota-Inclusion}
\end{equation}
\end{definition}

\noindent The first three constraints,
Eqs.~(\ref{eq:iota-constraints}), are the direct counterpart of the
corresponding QIVPM constraints, Eqs.~(\ref{eq:QIVPM-constraints});
the last condition, Eq.~(\ref{eq:iota-Inclusion}), is the direct
counterpart of the convexity conditions,
Eq{\color{red}s}.~(\ref{eq:classicalconvex}) and
(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
\cite{Choquet1954,Shapley1971,NgMoYeh1997,MarinacciMontrucchio2005}. Therefore,
these conditions guarantee that, for any conventional quantum
probability measure $\mu$, the composition $\ultramodular \circ \mu$
defines a valid QIVPM. Conversely, if for every quantum probability
measure $\mu$, it is the case that if $f \circ \mu$ is a QIVPM, then the {\color{blue}interval
map}~$f$ is an ultramodular function. Formally, we have the following
result{\color{red}:}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

We now give the necessary technical definitions to prove this
non-empty core property.
\begin{definition}[Consistency]\label{def:Consistency} We say a QIVPM
$\bar{\mu}$ is
\emph{consistent} with a state $\rho$ and a projector~$P$ if the
interval $\bar{\mu}(P)$ contains the exact probability calculated by
the Born rule~\cite{Born1983bibTeX,Mermin2007,Jaeger2007}, i.e.,
  \begin{equation}
  \muB_{\rho}\left(P\right) = \Tr\left(\rho
  P\right)\in\bar{\mu}\left(P\right)\ .\label{eq:consistent}
  \end{equation}
In contrast with classical probability spaces, there is \emph{no
guarantee} that there exists a unique state $\rho$ that satisfies
Eq.~(\ref{eq:consistent}) and therefore is consistent with a QIVPM.
\end{definition}

@@

{\color{red}An outline of t}he proof, detailed in the forthcoming
thesis~\cite{TaiThesis2018}, proceeds as
follows. From a subspace $\eventsC$ of mutually commuting events, one
can construct a partial orthonormal basis by diagonalization, and
complete this to a full orthonormal basis $\overline{\eventsC}$.  We
can then build a bijection between the QIVPM on the set of projectors
associated with this basis and the set of classical events
corresponding to this basis.  Using this correspondence together with
the classical result by
Shapley~\cite{Shapley1971,GilboaSchmeidler1994,NgMoYeh1997,Grabisch2016},
we can establish that for the special case of commuting events, a
QIVPM will always have a non-empty core.

@@

{\color{red}\subsection{Finite-Precision Extension of Gleason's Theorem}}

@@

{\color{red}\subsection{Ultramodular Functions}}

Theorem~\ref{thm:Finite-precision-Gleason} generalizes Gleason's
theorem in the sense that it accounts for a larger class of probability measures that includes
the conventional one as a limit. The theorem is however ``special'' in
the sense that it only applies to the particular class of QIVPMs
constructed by composing a{\color{red}n} interval map with a conventional
quantum probability measure. QIVPMs constructed in this manner have
some peculiar properties that we examine next.

A{\color{red}n} interval map is called \emph{ultramodular} if it satisfies
the following properties.

\begin{definition}[Ultramodular Functions]\label{def:THOS}Given
  a collection of intervals $\mathscr{I}$ including $\imposs$ and
  $\necess$, a{\color{red}n} interval map
  $\ultramodular:\left[0,1\right]\rightarrow\mathscr{I}$ is called
  ultramodular if:
\begin{subequations}\label{eq:iota-constraints}
\begin{eqnarray}
 &  & \ultramodular(0)=\imposs\,,\\
 &  & \ultramodular(1)=\necess\,,\\
 &  & \ultramodular\left(1-x\right)=\necess-\ultramodular\left(x\right)\,,
\end{eqnarray}
\end{subequations}and for any three numbers~$x_{0}$, $x_{1}$, and
$x_{2}\in\left[0,1\right]$ such that
$y=x_{0}+x_{1}+x_{2}\in\left[0,1\right]$, we have:
\begin{equation}
\ultramodular\left(y\right)+\ultramodular\left(x_{2}\right)\subseteq\ultramodular\left(x_{0}+x_{2}\right)+\ultramodular\left(x_{1}+x_{2}\right)\,.\label{eq:iota-Inclusion}
\end{equation}
\end{definition}

\noindent The first three constraints,
Eqs.~(\ref{eq:iota-constraints}), are the direct counterpart of the
corresponding QIVPM constraints, Eqs.~(\ref{eq:QIVPM-constraints});
the last condition, Eq.~(\ref{eq:iota-Inclusion}), is the direct
counterpart of the convexity conditions,
Eq{\color{red}s}.~(\ref{eq:classicalconvex}) and
(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
\cite{Choquet1954,Shapley1971,NgMoYeh1997,MarinacciMontrucchio2005}. Therefore,
these conditions guarantee that, for any conventional quantum
probability measure $\mu$, the composition $\ultramodular \circ \mu$
defines a valid QIVPM. Conversely, if for every quantum probability
measure $\mu$, it is the case that if $f \circ \mu$ is a QIVPM, then the {\color{red}interval
map}~$f$ is an ultramodular function. Formally, we have the following
result{\color{red}:}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{\color{red}\begin{acknowledgments}
The authors would like to thank the anonymous reviewers for their helpful and
constructive comments that greatly contributed to improving the final version of the paper.
\end{acknowledgments}}
